\chapter{Original Contributions}
\label{appendix:contributions}

Several papers were published in academic journals throughout the course of the doctoral studies, spanning from 2018 to 2022. Some of them are directly connected to the main topic of the thesis and others were published \emph{in parallel}, while working on other projects outside the topic of triaxial nuclei. As mentioned in the general conclusions presented in Chapter \ref{chapter-8-conclusions}, six scholarly articles are presented in this work and all of them were published in ISI indexed academic journals. Presentations and posters were also developed and shared at various conferences nation-wide or internationally. 

\section{Published Papers During Doctoral Studies}

Nomenclature:
\begin{itemize}
    \item \textbf{in-scope:} papers presented within this thesis, which were published between 2018-2022. Their results consists of original contribution on the field of Nuclear Structure. 
    \item \textbf{out-of-scope:} papers that were not presented in this thesis, which were published between 2018-2022. Their key results are not strictly related to the Ph.D. topic.
\end{itemize}

\subsection{In-Scope ISI Indexed Papers}

\begin{enumerate}
    \item Raduta, A. A., \textbf{R. Poenaru}, C. M. Raduta: \emph{New approach for the wobbling motion in the even-odd isotopes $^{161,163,165,167}$Lu}, Physical Review C, vol. 101, no. 1, \textbf{2020}, DOI: \href{https://link.aps.org/doi/10.1103/PhysRevC.101.014302}{10.1103/PhysRevC.101.014302}
    \item Raduta, A. A., \textbf{R. Poenaru}, C. M. Raduta: \emph{Towards a new semi-classical interpretation of the wobbling motion in $^{163}$Lu}, Journal of Physics G: Nuclear and Particle Physics 47(2), 025101, \textbf{2020}, DOI: \href{https://iopscience.iop.org/article/10.1088/1361-6471/ab5ae4/meta}{10.1088/1361-6471/ab5ae4}
    \item Raduta, A. A., C. M. Raduta, \textbf{R. Poenaru}: \emph{A new boson approach for the wobbling motion in even–odd nuclei}, Journal of Physics G: Nuclear and Particle Physics 48(1), 015106, \textbf{2020}, DOI: \href{https://iopscience.iop.org/article/10.1088/1361-6471/abc533/meta}{10.1088/1361-6471/abc533}
\end{enumerate}

\subsection{Out-of-Scope ISI Indexed Papers}

\subsection{Out-of-Scope Non-ISI Papers}

\section{Conferences}

\subsection{International}

\subsection{National}