\documentclass[a4paper, 12pt,openright,twoside]{Thesis}

\usepackage{amsmath}
\usepackage{mathtools}

% change the equation reference between parentheses
% \newtagform{brackets}{(}{)}
% \usetagform{brackets}

\usepackage{float}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{lipsum}
\usepackage{siunitx}
\usepackage{physics}
\usepackage{graphicx}
\usepackage{mathrsfs}
\usepackage{verbatim}  % Needed for the "comment" environment to make LaTeX comments
\usepackage{vector}  % Allows "\bvec{}" and "\buvec{}" for "blackboard" style bold vectors in maths
\usepackage[square, numbers, comma, sort&compress]{natbib}

\hypersetup{urlcolor=blue, colorlinks=true}  % Color hyperlinks in blue, but this can be distracting if there are many links.

%% --------------------------------- tikz ---------------------------------------
\usepackage{pgfplots}
\usepackage{tikz}
\pgfplotsset{compat=1.16}
\pgfplotsset{compat=1.16}
\usepgfplotslibrary{fillbetween}
\usetikzlibrary{cd,arrows,calc,fit,positioning,matrix,backgrounds}
\usetikzlibrary{shapes.geometric,shapes.multipart,decorations.pathreplacing}
\tikzstyle{startstop} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=white]
\tikzstyle{arrow1}=[ultra thick,->,>=stealth]
\tikzstyle{arrow}=[thick,->,>=stealth]
\tikzset{
    shift left/.style ={commutative diagrams/shift left={#1}},
    shift right/.style={commutative diagrams/shift right={#1}}
}

% define the gaussian function
\newcommand\gauss[2]{1/(#2*sqrt(2*pi))*exp(-((x-#1)^2)/(2*#2^2))} % Gauss function, parameters mu and sigma

%define proper imaginary unit within math-mode
\newcommand{\iu}{\mathrm{i}\mkern1mu}

% define the thesis title
\newcommand\mytitle{A Systematic Description of the Wobbling Motion in Odd-Mass Nuclei Within a Semi-Classical Formalism}

\bibliographystyle{unsrt}

%%-------------------------------------------------------MAIN DOCUMENT---------------------------------------------------%%
% ------------------------------------------------------------------------------------------------------------------------

\begin{document}

\frontmatter
\begin{titlepage}
    \begin{center}
        \textsc{ \LARGE{University of Bucharest \\}}
        \textsc{ \LARGE{Faculty of Physics\\ }}
        \vspace{40mm}
        \textup{\LARGE{\mytitle}}
    \end{center}
    
    \vspace{25mm}
    
    \begin{minipage}[t]{0.47\textwidth}
        \textnormal{\large{\bf Scientific Coordinator:\\}}
        {\large Prof. Dr. Em. A. A. Raduta}
    \end{minipage}\hfill\begin{minipage}[t]{0.47\textwidth}\raggedleft
        \textnormal{\large{\bf PhD Student:\\}}
        {\large Robert Poenaru}
    \end{minipage}
    
    \vspace{40mm}
    
    \centering{\large{A thesis submitted for the degree of \\ \emph{Doctor of Philosophy}}}
    
    \vspace{35mm}
    
    \centering{\large{\today}}
\end{titlepage}

\title{\mytitle}

\setstretch{1.3}  % Reset the line-spacing to 1.3 for body text (if it has changed)

\acknowledgements{
\addtocontents{toc}{\vspace{1em}}  % Add a gap in the Contents, for aesthetics
\emph{It has been quite a journey}... These five years showed me that many results can be achieved through hard work and dedication, but also much more can lie ahead of us in terms of new projects, new discoveries, and innovative theories.

Firstly, I would like to express my gratitude to Prof. A. A. Raduta, my coordinator and also mentor. His patience, work ethic, and constant support gave me the necessary motivation to keep going and improve in every aspect specific to academic research. Big thanks to my family for the tremendous emotional help and life advices offered when needed. Moreover, I would like to thank my highschool teacher Gabriela Pojoreanu, for igniting in me a love for Physics and fostering my curiosity about the laws of Nature.

I am forever grateful to my medical team - Dr. Demian, Dr. Mester, Dr. Radu and Dr. Zarma - whose unwavering support and expertise have been instrumental in my recovery and well-being. Completing my studies and this thesis would not have been possible without their attentive care. I want to extend heartfelt thanks to Aurica for her invaluable guidance in helping me and my family understand the importance of time in matters of health. And to my cousin Alex, who has been a constant source of support and encouragement - thank you for always being there.

Finally, I would like to believe that by reading this thesis our work and achieved results will be appreciated.
}
\clearpage

\tableofcontents
\pagestyle{fancy}  % Return the page headers back to the "fancy" style
\mainmatter

% Chapters
\input{Chapters/Chapter1} % Introduction
\input{Chapters/Chapter2} % Refactored Chapters 2 + 3 + 4
\input{Chapters/Chapter3}
\input{Chapters/Chapter4}
\input{Chapters/Chapter5}
\input{Chapters/Chapter6}
\input{Chapters/Chapter7} % General Conclusions

% Appendices 
\appendix
% Shell Model
\input{Chapters/Appendices/Appendix_ShellModel.tex}
% Bohr Model
% \input{Chapters/Appendix_Bohr.tex}
% Two-State Mixing
\input{Chapters/Appendices/Appendix_TwoState.tex}
% Rotational Bands
\input{Chapters/Appendices/Appendix_RotationalBands.tex}

% References
\bibliography{bibliography}

\end{document}
