\chapter{Bohr's Collective Model}
\label{appendix:bohr-model}

calculations in Refs. \cite{ring2004nuclear,bertulani2007nuclear}) and 

Indeed, by looking at the expression from Eq. \ref{collective-hamiltonian-stiffness-inertia}, one can see that it can be brought to a form:
\begin{align}
    H_\text{coll}^\text{osc}=\frac{p^2}{2m}+\frac{1}{2}kr^2\ ,
    \label{eq-bohr-hamiltonian-oscillator-simple}
\end{align}
which is typical to a harmonic oscillator Hamiltonian. The frequency of oscillation is given by the relation $\omega=\sqrt{k/m}$. The vibrations can now be understood in terms of a sum of harmonic oscillator frequencies, where each frequency is given by $\lambda$:
\begin{align}
\omega_\lambda=\sqrt{\frac{C_\lambda}{B_\lambda}}\ .
\end{align}

The inertia term $B_\lambda$ (also called the \emph{mass parameter}) has the following expression \cite{ring2004nuclear}:
\begin{align}
    B_\lambda=\rho \frac{mR_0^5}{\lambda}=\frac{3}{4\pi\lambda}AmR_0^2\ ,
    \label{inertia-parameters-B}
\end{align}
showing a quadratic dependence with the average nuclear radius. The nuclear density of nucleons with mass $m$ is given by $\rho$. The stiffness parameter $C_\lambda$ is usually expressed in terms of the \emph{surface tension} $\sigma$, assuming that the nuclear matter exhibits an irrotational flow and the nuclear charge is distributed uniformly over the entire volume \cite{ring2004nuclear}:
\begin{align}
    C_\lambda=(\lambda-1)(\lambda+2)\sigma R_0^2-\frac{3}{2\pi}\frac{\lambda-1}{2\lambda+1}\frac{(Ze)^2}{R_0}\ .
    \label{stiffness-parameters-C}
\end{align}

With these ingredients, one can sketch a collective Hamiltonian similar to Eq. \ref{eq-bohr-hamiltonian-oscillator-simple} in the following manner:
\begin{align}
    H_\text{coll}=\sum_{\lambda\mu}\hbar\omega_\lambda\left(N_{\lambda\mu}+\frac{1}{2}\right)\ ,
\end{align}
where indeed, the typical harmonic oscillator spectrum emerges. The recipe for further manipulation of the Bohr's collective Hamiltonian from Eq. \ref{collective-hamiltonian-stiffness-inertia} is a lengthy process \cite{bohr1998nuclear,ring2004nuclear}, with several considerations beyond the scope of the current work. Shortly, for the quadrupole deformed nuclei, the expression of the potential term $V$ will be a function that depends on the parameters $(\beta,\gamma)$. The potential will be defined as a quadratic approximation in the vicinity of a deformed minimum point $p_0|_\text{min}=(\beta_,\gamma_0)$. This starts from the general assumption that the nucleus has the deformation $p_0$ in the ground state, and excitations around the \emph{equilibrium} point. Thus $V(\beta,\gamma)$ can be written as:
\begin{align}
    V(\beta,\gamma)=\frac{1}{2}C_{0}\left(\alpha_{20}(\beta,\gamma)-\alpha_{20}^0\right)^2+\frac{1}{2}C_{2}\left(\alpha_{22}(\beta,\gamma)-\alpha_{22}^0\right)\ .
    \label{bohr-collective-potential}
\end{align}